---
title: "Detection of off-bottom movements from Monkfish DST data"
output: rmarkdown::github_document
---

Extensive vertical movements of monkfish can be recorded by the data storage tags (DSTs). Such behavior may be related to spawning, foraging, or tidal-based migration behavior. In this notebook I explore a method to detect off-bottom movements from monkfish DST data. I first remove the baseline from the depth timeseries data and then detect peaks in the corrected signal.

```{r}
library(R.matlab)
library('baseline')
library("wmtsa")
library(ggplot2)
source('matlab_time.R')
```

Read the preprocessed tag data:
```{r}
matlabFile  <- readMat('/Users/cliu/Dropbox/Geolocation/projects/monkfish/Preprocessing/processed_tags/10_raw.mat')
varNames    <- names(matlabFile$tag[,,1])
tag     <- matlabFile$tag
tag     <- lapply(tag, unlist, use.names=FALSE)
names(tag) <- varNames
```

Plot the depth timeseries:
```{r}

df<-data.frame(date=matlab2POS(tag$dnum), depth=tag$depth)
data <- split(df$depth, as.Date(df$date))
plot(df,type="l",ylim = rev(range(tag[["depth"]])), col="blue")
```

Use the `baseline` package to detece the baseline signal, and correct the signal (substracting the baseline from the original signal):
```{r}
peakDetection <- baseline(t(-df$depth), method = "peakDetection", left=36, lwin=5)
plot(peakDetection, labels=df$date)
```

Use the `wmtsa` package to detect peaks in the corrected signal:
```{r}
x = wavCWT(t(peakDetection@corrected))
xtree = wavCWTTree(x)
P = wavCWTPeaks(xtree, noise.min=7, noise.span=36)
plot(t(peakDetection@corrected), type="l")
points(P$x, P$y,pch=3,col="red")

plot(-df$depth, type="l")
points(P$x, -df$depth[P$x],pch=3,col="red")
lines(t(peakDetection@baseline), type="l", col="blue")
```

```{r}
Pdf = data.frame(x=df$date[P$x], y=df$depth[P$x])
ggplot(data=df, aes(x=date, y=depth)) +geom_line() +geom_point(data=Pdf, aes(x=x, y=y), shape=3, color="red") +scale_y_reverse() 
```

Output the detected off-bottom movement events into a dataframe:
```{r}
df_offbottom <- df[P$x,]
rownames(df_offbottom) <- 1:nrow(df_offbottom)
df_offbottom
```


